Log or not - Analysis of ECDC forecasts

### 2022-11-21 analysis/Analysis ECDC forecasts.Rmd
update correlation plot, add plot with number of available forecasts

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)
library(tidytext)
library(grid)

@@

@@

hub_data <- hub_data[model %in% filtermodels]
hub_data[, true_value := pmax(true_value, 0)]

@@

General thoughts?

- Focus on Hub-ensemble or average across all models? 
- Focus on two weeks ahead?

@@

@@

## Figure 5 - Correlations between everything
```{r}
df <- scores |>
  filter(horizon < 5) |>
  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction, type_and_scale)) |>
  pivot_wider(names_from = scale, values_from = interval_score) |> 
  group_by(horizon, target_type, forecast_date, target_end_date, location) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman")) |>
  group_by(horizon, target_type) |>
  summarise(cor = mean(cor, na.rm = TRUE), 
            spearman = mean(cor, na.rm = TRUE)) |>
  rename(`Target type` = target_type)

@@

(scatter_wis_obs) / (p_cor_scores + p_cor_skill) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(1, 1), heights = c(2, 1)) &
  theme(legend.position = "bottom")

@@

>>>>

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)
library(tidytext)
library(grid)
library(ggpubr)

@@

remove <- hub_data |> 
  group_by(model, forecast_date, location, target_type) |>
  summarise(q = length(unique(quantile))) |>
  filter(q < 23) |>
  pull(model) |>
  unique()

filtermodels <- setdiff(filtermodels, remove)

@@

hub_data <- hub_data[model %in% filtermodels]
hub_data <- hub_data[forecast_date >= "2021-03-05"]
hub_data[, true_value := pmax(true_value, 0)]

@@

## Basic data summary
```{r}
hub_data |>
  summarise(min_date = min(forecast_date), 
            max_date = max(forecast_date)) |>
  mutate(week_diff = (max_date - min_date - 1) / 7)

hub_data$location |> 
  unique() |>
  length()

hub_data |> 
  filter(horizon == 2) |>
  avail_forecasts(by = c("target_end_date", "target_type", "model")) |>
  mutate(forecast_date = as.Date(target_end_date) - 12) |>
  plot_avail_forecasts(show_numbers = FALSE) +
  facet_grid(~ target_type) + 
  labs(y = "Model", x = "Forecast date")

ggsave(filename = "output/figures/number-avail-forecasts.png", height = 4, width = 10)
```

@@

## Figure 5 - Correlations between everything
```{r}
wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {

@@

scatter_wis_obs <- wis_vs_total(scores) + 
  labs(y = "Mean WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  ggpubr::stat_cor(method = "spearman", p.accuracy = 0.001, 
                   cor.coef.name = "rho", size = 3)

df <- scores |>
  filter(horizon < 5) |>
  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction, type_and_scale)) |>
  pivot_wider(names_from = scale, values_from = interval_score) |> 
  group_by(horizon, target_type, forecast_date, target_end_date, location) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman")) |>
  group_by(horizon, target_type) |>
  summarise(cor = mean(cor, na.rm = TRUE), 
            spearman = mean(cor, na.rm = TRUE)) |>
  rename(`Target type` = target_type)

@@

(scatter_wis_obs) / (p_cor_scores + p_cor_skill) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(1, 1), heights = c(3, 1)) &
  theme(legend.position = "bottom")

@@

```{r}
s <- hub_data |>
  filter(model %in% c("EuroCOVIDhub-ensemble", "epiforecasts-EpiNow2")) |>
  score()

s_summarised <- s |>
  summarise_scores(by = c("model", "range", "horizon"), metric = c("interval_score"))

s_summarised |>
  select(c(model, interval_score, range, horizon)) |>
  pivot_wider(names_from = model, values_from = interval_score) |>
  mutate(rel = `epiforecasts-EpiNow2` / `EuroCOVIDhub-ensemble`) |>
  filter(horizon == 4)
  ggplot(aes(y = rel, x = range, 
             color = factor(horizon), group = factor(horizon))) +
  geom_line()

s_summarised |>
  filter(horizon == 1) |>
  select(c(model, interval_score, range, horizon)) |>
  ggplot(aes(y = interval_score, x = range, 
             color = model, group = model)) +
  geom_line()
